Pragmatic interactions lead to efficient language structure and use

### 2019-01-21 paper/paper.Rmd
minor clean-up improvments and clarifications, removed conditional efficiency plots / section for space and narrative flow.

Why do languages look they way they do? Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization. In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon. Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent properties of speaker-listener effort minimization. \par

@@

Zipf’s original work and subsequent functionalist projects describe a varied set of linguistic properties and human behaviors. Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication. Simply put, what is “hard” for a speaker is likely different from what is “hard” for a listener. Zipf characterized this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ should map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well. The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena. Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiency we see in languages is fundamentally related to local-level efficiency during the in-the-moment interactions of rational, pragmatic speakers and listeners at the level of conversation. \par

In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation. To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain, (2) a framework for describing in-the-moment interactions of speaker-listeners, and (3) some measure of linguistic efficiency. For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi et al. (2012) into our framework. To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016). Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting. In terms of our opening question, we are exploring the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”? In more general terms, this project is meant to demonstrate the ways in which a range of functionalist theories might make use of such a framework -- e.g. we could replace \textit{lexical ambiguity} with a range of behaviors and properties identified by functionalist approaches to language science.\par

We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games. Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting. We move on to two case-studies examining the questions posed above, examining the efficiency of ambiguity from a language design and language use perspective.

@@

```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology 'language')."}
img <- png::readPNG("figs/game1.png")
grid::grid.raster(img)
```

### Reference games
Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein). We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this setting, speakers and listeners are aware of a set of objects $M$, which will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that could be used to refer to a given meaning. Utterances may have different relative costs, operationalized via a prior over utterances $p(u)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $p(m)$. Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION). We consider a set of contexts $C$ which describe different need probability states, formalized as different conditional distributions over meanings $p(m|c)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, referent pairs.

@@

We assume that the particular joint distribution over utterance-meaning $<u, m> = e$ pairs follows from a simple generative model. First, some meaning is sampled with probability $p(m)$. Our speaker attempts to convey this intended meaning to a listener by encoding it in the utterance $u$ via the conditional $S(u|m; \ell)$. Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = P_{speaker}(u, m; \ell) = S(u | m; \ell)p(m)$.

@@

Piantadosi et al. (2012) argue just the opposite. They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*. Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low. Context can provide useful information for disambiguation.\par

@@

The argument outlined by Piantadosi et al. (2012) is compatible with our current reference game setting with just a single addition -- context. The authors argue that mapping both $m_1$ and $m_2$ to $l_1$ is useful when they can be disambiguated *in context*. In our case we consider a context $c$ to specify a unique ordering of the need probabilities. That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.

@@

Note that in the case that $|C| =1$, our objective simplifies to our original equation (7). With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (e.g. the size of $|C|$).\par

@@

Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents, as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$). Rather than assume particular structure for our utterance and meaning prior distributions, for each simulation we generate $P(M) \sim \text{Dir}(1, |M|)$, $P(U) \sim \text{Dir}(1, |U|)$ and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.

#### Context
Following the argument given by Piantadosi et al. (2012) we want to assess the impact of *context* on our objective. To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$. Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $p(m|c_1)$. Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $p(m|c_1)$ as well as $p(m|c_2)$. The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.

#### Baselines
For comparison, we also examine properties of optimal languages under two additional objectives. Zipf (1935) proposed that the optimal speaker language should consist of a single low cost word $\ell_{speaker}^*$. We operationalize this using the *first half* of equation (1).

@@

The optimal listener language $\ell_{listener}^*$, by contrast, should bijectively map all words to unique meanings. We operationalize this using the *second half* of equation (1).

@@

### Results
Figure 1. plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener}; \ell)$ for each context condition. The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell*$ under our speaker-listener cross-entropy objective contains ambiguity. We also plot the proportion of optimal speaker-only and listener-only languages that contain ambiguity. In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single utterance (full ambiguous). Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided (bijective form-meaning mappings).\par

#### More context leads to larger efficiency gains
Piantadosi et al. (2012) framed their theory in terms of properties of conditional entropy. That is,  $H(X|C) \leq H(X)$. This inequality is strict when $C$ provides information about $X$ (Cover & Thomas, XXXX). Intuitively, this means that they amount of uncertainty we have about $X$ decreases when $C$ tells us something about $X$. In our setting, knowing the particular context $c_i$ is always informative with respect the distribution over meanings ($p(m|c_i)$). This means that if we have $|C_1|=1$ and $|C_2|=2$ then we expect $H_{cross}(P_{Speaker}, P_{Listener}|C_2) < H_{cross}(P_{Speaker}, P_{Listener}|C_1)$. To examine this in our simulations, we look at the difference in efficiency scores, comparing pragmatic agents who have access the conditional distributions $P(m|c)$ an non-pragmatic agents who don’t and simply consider $P(m) =

Additionally, the difference between these two quantities should increase with the difference in the size of $|C_i|$. E.g. if we also have $|C_3|=3$ then  . In our case, we can examine the difference between the unconditional and conditional cross-entropy measures as $|C|$ increases from $|C|=1$ in condition 1 to $|C|=4$ in condition 4. The right facet of figure one shows the difference between conditional and unconditional corr-entropy measures as the number of contexts $|C|$ increases.

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Left facet -- optimal languages are more likely to contain ambiguous items as amount of contextual information increases. Vertical axis shows the proportion of optimal languages that contain ambiguity. Horizontal axis shows the number of contexts in each condition (1-4). Red-line represents the optimal langauge under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration. Right facet -- Gains in efficiency increase with the amount of contextual information (as the number of contexts increases). Vertical axis shows the difference in cross-entropy when the number of contexts is 1 and when the number of contexts is greater than 1. Horizontal axis shows each of the four conditions (the four context sizes). Note that the difference for condition 1 is zero as there is only a single context."}

@@

### Summary
Piantadosi et al. (2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context. Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative. Further, we showed that the impact of being able to disambiguate language (having access to $p(m|c_i)$) is increasingly efficient as the amount of common-ground (context) increases.\par

@@

Our first experiment made a fairly direct test of the communicative function of ambiguity proposed theory presented by Piantadosi et al. (2012). Sampling a set of need probabilities and utterance costs, we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective. Results indicated that ambiguity is an efficient property when context is informative. This simulation, however, assumed that our speaker and listener agents could always disambiguate items context. That is, in our four conditions both agents had perfect knowledge of the conditional distributions $p(m|c_i)$.\par
However, in day to day communication we often don’t have perfect knowledge of the current context of a conversation. We can consider this notion of “context” as analogous to knowledge of the *topic* of conversation. Consider the scenario in which a reader is beginning a newspaper article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context to disambiguate it. Using such ambiguity is only warranted when the appropriate person or event is salient in the context.\par

We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse. In this scenario, we assume a single language $\ell$ which contains ambiguous items, known to both the speaker and listener. Crucially, we do not assume that a listener has access the conditional distributions $p(m|c)$ a priori. Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but does not know which one is currently being used by the speaker.\par

Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c$, as well as the particular meaning of a given utterance. That is we consider a listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$. Note that both the speaker and listener can track the history of previous utterances $D$.\par

@@

### Hypothesis
In this setting, if speakers are *using ambiguity efficiently* they should only use the ambiguous material when they know the listener can disambiguate the items in context. Since listeners accumulate information about the current topic distribution ($p(m|t)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.

@@

We conduct $n=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:

@@

We assume that $p(u_5) = p(u_6) > p(u_{1,...,4})$. That is, the two ambiguous utterances are less costly than the non-ambiguous utterances. In this setting, it is less costly for the speaker to use the ambiguous utterances. However, if the listener cannot disambiguate them, then they will incur a disambiguation cose. We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$. This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)$. In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the context $p(m|C=c)$.

@@

The right facet of figure 2 shows the listener belief about the current topic $c \in C$ given the discourse history ($p(c|D)$). Note that the trajectory of these items are close -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material. Figure 4 shows the listener belief about the current context. Additionally, we can track the behavior of our cross-entropy objective and it’s component speaker and listener elements over the discourse. Overall we see declines in the speaker-listener cross-entropy. These are derived primarily from the listener learning about the current context of conversation (i.e. the current $c$ defining $p(m|c)$), however we also see declines in speaker effort, as they are able to use loss-costly material later in the discourse.\par

@@

These simulations demonstrated the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al. (2012) -- Only once the context is known to the listener does it makes sense to use the lower-cost material.

@@

How might the competing pressures of speakers and listeners give rise to the distributional forms found in natural language? Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon. Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency. But, it is not clear at what level this optimization occurs. In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment local interactions between speakers and listeners. We explore this idea in our current project, following theoretical work by Lawrence Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975). To so, we explored the degree to which the interactions of rational pragmatic agents can provide a framework for understanding efficient language structure and use. We focused on an argument on the communicative function of ambiguity put forth by Piantadosi. We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative. Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context. The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework. Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS) and  the emergence of compositionality (CITATIONS), as well as efficient examples of language use including the efficient use of reduction and redundancy at the level  of syntax (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).

# References

>>>>

Why do languages look they way they do? Zipf (1935) proposed that distributional properties found in natural language were evidence of speaker-listener effort minimization. In his own words, “we are arguing that people do in fact act with a maximum economy of effort, and that therefore in the process of speaking-listening they will automatically minimize the expenditure of effort.” Evidence for this claim was largely derived at the level of the lexicon. Zipf argued that the particular relationship between a word’s frequency and its rank, length, and denotation size could be explained as an emergent property of speaker-listener effort minimization. \par

@@

Zipf’s original work and subsequent functionalist projects describe a varied set of linguistic properties and human behaviors. Common to many is a characterization of a fundamental effort-asymmetry underlying everyday communication. Simply put, what is “hard” for a speaker is likely different from what is “hard” for a listener. Zipf characterized this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of just a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

This move to study efficient language use in terms of reference games has clear theoretical underpinnings as well. The linguist Lawrence Horn (1984), highlighted the importance of Zipf’s principles for explaining conversation-level phenomena. Horn suggested a link between Zipf’s conception of speaker- and listener-effort and ideas related to conversational pragmatics by Grice (1975), arguing that the interaction of Zipf’s forces were “largely responsible for generating Grice’s conversational maxims and the schema for pragmatic inference derived therefrom.” Put differently, the system-level efficiencies we see in language are fundamentally related to the local-level efficiencies that occur during rational, pragmatic conversation.\par

In this work, we present a step toward formalizing the framing presented above  -- connecting Zipfian notions of efficient language structure to Gricean notions of rational conversation. To do so minimally requires three basic ingredients -- (1) a language property we’d like to explain, (2) a framework for describing in-the-moment interactions of speaker-listeners, and (3) some measure of linguistic efficiency. For the current project we focus on the “communicative function of ambiguity” for our property -- translating the functionalist theory pursued by Piantadosi et al. (2012) into our framework. To model local, conversational interactions (2) we adopt the Rational Speech-act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016). Finally, (3) we derive an objective function for measuring language efficiency in the reference game setting.\par

The current work examines the question, “how does having \textit{lexical ambiguity} make using language $\ell$ more or less useful for communication”? We begin with a high-level introduction to the modelling framework introducing the basic ingredients we will need to represent language as repeated reference games. Following this introduction we derive a simple objective function for measuring the efficiency of linguistic systems in this setting. We move on to two case-studies examining the questions posed above, examining the efficiency of ambiguity from a language design and language use perspective.

@@

```{r plot-reference-game, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2.5, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.cap = "An example reference game with associated literal semantics (in our terminology a ‘language’)."}
img <- png::readPNG("figs/game1.png")
grid::grid.raster(img)
```

### Reference games
Zipf’s canonical (1935) example of optimal speaker- and listener-languages took the form of a reference game (similar to those described by Wittgenstein). We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this setting, speakers and listeners are aware of a set of objects $M$, which we will refer to as *meanings* and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning. Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$. Note that the prior over meanings are analogous to the *need probabilities* assumed in previous work (Regier, CITATION). We consider a set of contexts $C$ with an associated prior $P(C)$. Importantly, contexts describe different need probability states, formalized as different conditional distributions over meanings e.g. $p(M|C=c_i)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m> = e$ is a tuple of utterance, referent pairs.

@@

We assume that the particular joint distribution over utterance-meaning $<u, m> = e$ pairs follows from a simple generative model. First, some meaning is sampled with probability $p(m)$. Our speaker attempts to convey this intended meaning to a listener via an utterance $u$ by sampling from the speaker conditional distribution $S(u|m; \ell)$. Combining these terms leads to the *speaker's joint distribution over events* which we can write as $P(e) = S(u | m; \ell)p(m) = P_{speaker}(u, m; \ell)$.

@@

Piantadosi et al. (2012) argue just the opposite. They claim that ambiguity is an *efficient* property of any communication system in which *communication is contextualized*. Simply put, it is useful to have a language that re-uses low-cost material (has ambiguity) so long as the cost of disambiguating the material is low. In particular, context (or common ground) can provide useful information for disambiguation. Say we have two objects in the world $m_1$ and $m_2$ and two languages $\ell_1$ and $\ell_2$. In language $\ell_1$ the low-cost utterance $u_1$ can be used to refer to both $m1$ and $m_2$ (i.e. $[[u_1]]_{\ell_1} = \{m_1, m_2 \}$. Language $\ell_2$, by contrast, defines a mapping such that the low cost utterance $[[u_1]]_{\ell_2} = \{m_1\}$ and a higher cost utterance $[[u_2]]_{\ell_2} = \{m_2 \}$. While it is cheaper for a speaker to use $\ell_1$ (because she can always use a low-cost utterance), it is more difficult for a listener (because $u_1$ is ambiguous). However, if the meaning $u_1$ can be disambiguated *in context* then the listener will not incur disambiguation cost and the speaker can always use the low-cost utterance. This a basic argument for the contextual efficiency of polysemy. If I use the word “bat” at a baseball game a listener will likely infer I am referring to the baseball sense of the word. If I use the word “bat” in a dark cave, a listener will likely infer I am referring to the animal.

@@

The argument outlined by Piantadosi et al. (2012) is compatible with our current reference game setting with just a single addition -- context. The authors argue that mapping two meanings $m_1$ and $m_2$ to a single utterance $u_1$ is useful when they can be disambiguated *in context*. In our case we consider a context $c$ to specify a unique ordering of the need probabilities. That is, $p(m|c_1) \neq p(m|c_2)$, when there are two contexts $|C| = \{c_1, c_2\}$.

@@

Note that in the case that $|C| =1$, our objective simplifies to our original equation (7). With this update we can compare languages $L$ while varying the degree to which context contains useful information, varying the number of contexts (the size of $|C|$).\par

@@

Recall that language efficiency is both a function of the particular semantic mappings induced by that language, the speaker and listener agents ($S$ and $L$), as well as the utterance ($P(U)$), meaning ($P(M)$), and context priors ($P(C)$). Rather than assume particular structure for our utterance, meaning and context prior distributions, for each simulation we generate $P(U) \sim \text{Dir}(1, |U|)$, $P(M) \sim \text{Dir}(1, |M|)$, and $P(C) \sim \text{Dir}(1, |C|)$ where $\text{Dir}(1, k)$ specifies the uniform Dirichlet distribution over a $k$-dimensional probability vector.

#### Context
Following the argument given by Piantadosi et al. (2012) we want to assess the impact of *context* on our objective. To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M), P(C)\}$. Our first is a *single-context* condition ( $|C|=1$) -- there is a only a single context describing $P(M)$. Our second condition contains two-contexts  ($|C| = 2$) -- we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$. The third and fourth condition correspond accordingly with $|C|=3$ and $|C| = 4$, respectively.

#### Baselines
For comparison, we also examine properties of optimal languages under two additional objectives. Zipf (1935) proposed that the optimal speaker language $\ell_{speaker}^*$ should only optimize speaker effort. We operationalize this using the *first half* of equation (1).

@@

The optimal listener language $\ell_{listener}^*$, by contrast, should only optimize listener effort. We operationalize this using the *second half* of equation (1).

@@

### Results
Figure 1. plots the proportion of optimal languages under $H_{cross}(P_{Speaker}, P_{Listener}; \ell)$ for each context condition. The red line indicates that as the number of contexts increases, so does the probability that the optimal language $\ell^*_{cross}$ under our speaker-listener cross-entropy objective contains ambiguity. We also plot the proportion of optimal speaker-only $\ell^*_{speaker}$ and listener-only $\ell^*_{listener}$ languages that contain ambiguity. In line, with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages will assign all meanings to a single utterance (full ambiguous). Likewise, if languages are designed only to minimize speaker listener effort then ambiguity should always be avoided (bijective form-meaning mappings).\par

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Optimal languages are more likely to contain ambiguous items as amount of contextual information increases. Vertical axis shows the proportion of optimal languages that contain ambiguity. Horizontal axis shows the number of contexts in each condition (1-4). Red-line represents the optimal language under our Zipfian cross-entropy objective while the blue and red lines show optimal languages under speaker- and listener-only consideration."}

@@

### Summary
Piantadosi et al. (2012) argued that it is useful to re-use low-cost linguistic forms for multiple meanings when they can be disambiguated in context. We simulated this effect in a reference game setting, examining the probability that an optimal language will contain an word with an ambiguous mapping as a function of context size. Using our speaker-listener cross-entropy measure of efficiency we showed that optimal languages are more likely to have ambiguous items when context is informative, but not under objectives which emphasizes speaker- or listener-effort alone.\par

@@

Our first experiment made a fairly direct test of the communicative function of ambiguity theory presented by Piantadosi et al. (2012). Sampling a set of need probabilities and utterance costs, we explored the space of possible languages, examining the properties of the language $\ell^*$ which minimized our objective. Results indicated that ambiguity is an efficient property when context is informative. This simulation, however, assumed that our speaker and listener agents could always disambiguate items context. That is, in our four conditions both agents had perfect knowledge of the conditional distributions $P(M|C_i)$.\par

This is a strong assumption and likely unrealistic for much of day-to-day communication.  In talking with another person, we often \textit{don’t have perfect knowledge of the current context of a conversation}. We can consider this notion of “context” as analogous to a *topic* of conversation. Consider the scenario in which a reader is beginning a newspaper article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this scenario, using an ambiguous pronoun early in the article lead to misunderstanding as there isn’t sufficient context for disambiguation. Using such ambiguity is only warranted when the appropriate person or event is salient in the context. \par

We consider such a scenario in simulation two -- analyzing efficient *use* of ambiguity over discourse. In this scenario, we assume a single language $\ell$ which contains both ambiguous and unambiguous utterances. The ambiguous utterances are less costly than the unambiguous utterances. Crucially, we do not assume that a listener has access the conditional distributions $p(M|C=c)$ a priori. Rather, we assume that the listener has knowledge of the set of possible contexts $C = \{c_1, \dots, c_k\}$, but \textit{does not know which one is currently being used by the speaker}. In this setting, a speaker minimizing her production effort would like to use the low-cost material as frequently as possible. However, this will incur a disambiguation cost to the listener if they are uncertain of the current context $c_{current}$ (topic of conversation).\par

Over the course of a discourse, $D$ the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$. That is we consider a listener model $L(m, c|u,D;\ell)$ and speaker model $S(u|m,c,D;\ell)$. Note that both the speaker and listener can track the history of previous utterances $D$. Importantly, a listener can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par

@@

### Hypothesis
In this setting, if speakers are *using ambiguity efficiently* they should only use the ambiguous material when they know the listener can disambiguate the items in context. Since listeners accumulate information about the current topic distribution ($P(C=c_{current}|D)$) over as the discourse progresses, we should expect to see more frequent use of ambiguous items later in the discourse.

@@

We conduct $N=200$ simulations, generating discourses of length $|D|=30$ utterances.  We consider a single language $\ell$ with $|U|=6$ and $|M|=4$ specified by the boolean matrix:

@@

We assume that $p(u_5) = p(u_6) > p(u_{1})=\dots=p(u_{4})$. That is, the two ambiguous utterances are less costly than the non-ambiguous utterances. In this setting, it is less costly for the speaker to use the ambiguous utterances. However, if the listener cannot disambiguate them, then they will incur a disambiguation cose. We assume that over the discourse both agents have complete recall of the set of possible utterances by the speaker $D$. This means that the listener can infer the current topic distribution $p(C=c|D)$ via $p(C=c|D) \propto \prod_{i=0}^{|D|}S(u_i|m_i)p(m_i|c)p(c)$. In this way, an efficient speaker should avoid ambiguity \textit{early} in the discourse when the listener is unsure of the context, only using it later once they are confident the listener knows the current context $p(C=c_{current}|D)$.

@@

The right facet of figure 2 shows the listener belief about the current topic $c_{current} \in C$ given the discourse history ($p(C=c_{current}|D)$). Note that the trajectories in plot A and B are related -- the speaker is more likely to use ambiguous material only when the listener is able to disambiguate that material.

@@

These simulations explored the efficient *use* of ambiguous material over a discourse, in an alternative instantiation of the argument outlined by Piantadosi et al. (2012). In the first set of simulations we varied the set of languages, examining the properties of languages which minimized our cross-entropy objective under different context conditions. In the current simulation we assumed that the listener did not know the current context (toptic) a priori, but had to infer it from the discourse. We examined speaker production behavior, tracking use of low-cost, but ambiguous material over the discourse. We found that early in the discourse our rational speaker avoided using ambiguous material, only using the material later when the context was known to the listener.

@@

How do the competing pressures of speakers and listeners give rise to the distributional forms found in natural language? Zipf (1935) proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon. Subsequent functionalist approaches to language science have provided additional evidence that various aspects of language structure and language use appear to be optimized for efficiency. In Zipf’s famous example of an optimal speaker and listener language he appears indicate that such optimization occurs in in-the-moment, local interactions between speakers and listeners. We explore this idea in our current project, following theoretical work by Horn (1984) who proposed the speaker-listener pressures identified by Zipf can also be interpreted in terms of the maxims governing rational conversation proposed by Grice (1975). To so, we explored the degree to which the interactions of rational pragmatic agents can provide a framework for understanding efficient language structure and use. We focused on an argument on the communicative function of ambiguity put forth by Piantadosi et al. (2012). We derive a novel speaker-listener cross-entropy objective for measuring the efficiency of a language in a reference game setting showing that optimal languages are more likely to contain ambiguous material when context is informative. Further we show that rational pragmatic agents will *use* ambiguous material efficiently, only when such use is supported by context. The intent is that this project provides a demonstration for the types of functionalist theories that can be explored within this framework. Among other language properties open to this type of analysis are work examining semantic typology (CITATIONS), the emergence of compositionality (CITATIONS) and convention formation, as well as efficient examples of language use including the efficient use of reduction and redundancy (Levy & Jaeger, 2007) and discourse (Genzel & Charniak, 2002).

# References 
Chomsky (2002)
Cover & Thomas, XXXX)
Ferrer-i-cancho, et al.,  201X
Frank & Goodman (2012)
Genzel & Charniak (2002)
Goodman & Frank (2016)
Grice (1957)
Horn (1984)
Kibry, Smith & Folds (201X)
Levy & Jaeger (2007)
Piantadosi et al. (2012)
Regier, et al.
Wasow, Perfors & Beaver (chomsky201X)
Zipf (1935)
